---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dpi = 600
)
```

# sportyR

<!-- badges: start -->
[![Travis build status](https://travis-ci.com/rossdrucker/sportyR.svg?branch=master)](https://travis-ci.com/rossdrucker/sportyR)
[![R-CMD-check](https://github.com/rossdrucker/sportyR/workflows/R-CMD-check/badge.svg)](https://github.com/rossdrucker/sportyR/actions)
[![License: MIT](https://img.shields.io/badge/License-MIT-orange.svg)](https://opensource.org/licenses/MIT)
[![star this repo](https://githubbadges.com/star.svg?user=rossdrucker&repo=sportyR&style=flat)](https://github.com/rossdrucker/sportyR)
[![fork this repo](https://githubbadges.com/fork.svg?user=rossdrucker&repo=sportyR&style=flat)](https://github.com/rossdrucker/sportyR/fork)
<!-- badges: end -->

As the field of sports analytics evolve, there's a growing need for methods to both track and visualize players throughout the game. This package aims to make this easy regardless of sport needed to be plotted.

## Installation

You can install the released version of `sportyR` from [GitHub]() with:

``` r
# install.packages("devtools")
devtools::install_github("rossdrucker/sportyR")
```

Once the library is installed, be sure to load it into the working environment.

```{r load-sportyR}
library(sportyR)
```


## Plotting Functions

All plotting functions in this library are named as `geom_{sport}()`, and take most (if not all) of the following arguments:

 - `league`: the league code for the sport. In all functions, this will ***NOT*** have a default value. The supplied league is **case-insensitive**. Future iterations of this package may allow the full league name to be supplied if desired (e.g. `league = 'National Basketball Associaton'` instead of `league = 'NBA'`), but this feature is not currently available.
 
 - `full_surf`: a boolean indicating whether or not to plot the full surface. This defaults to `TRUE`. <br>
   ***NOTE**: this argument is not taken in `geom_baseball()`, `geom_football()`, or `geom_soccer()` as these surfaces are always shown in full.*
 
 - `rotate`: a boolean indicating whether or not to rotate the surface. All surfaces will be horizontal by nature. This defaults to `FALSE`
 
 - `rotation_dir`: the direction in which to rotate the plot (if `rotate == TRUE`). This defaults to `'ccw'` for counterclockwise.

## Helpful Functions

In addition to the plotting functions, there are functions that are available to help the user see what is possible to be plotted. These are the `caniplot` functions: `caniplot_league()` and `caniplot_sport()`. Call `?caniplot_sport` and `?caniplot_league` for more information on these functions, but an example is shown below.

```{r caniplot-example}
caniplot_league('mlb')
```

## Examples

Most playing surfaces are standard in size, so they can be rendered via a call to the proper `geom_{sport}()` like so:

```{r nhl-example}
geom_hockey('nhl')
```

However, certain functions are able to take additional parameters. As an example, soccer pitches are not always a standard size. For this reason, users can specify in the call to `geom_soccer()` what the touchline and goal line dimensions should be (in meters).

```{r fifa-example}
geom_soccer('fifa', touchline_length = 100, goal_line_length = 75)
```

It's also possible to plot half-surfaces and rotated surfaces:

```{r ncaa_bb-example}
geom_basketball('ncaa', full_surf = FALSE, rotate = TRUE)
```

## Authorship

This package was developed and is being maintained by [Ross Drucker](https://github.com/rossdrucker). Please reach out with any questions, bugs, or suggestions.
