---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sportyR

<!-- badges: start -->
<!-- badges: end -->

As the field of sports analytics evolve, there's a growing need for methods to both track and visualize players throughout the game. This package aims to make this easy regardless of sport needed to be plotted.

## Installation

You can install the released version of `sportyR` from [GitHub]() with:

``` r
# install.packages("devtools")
devtools::install_github("rossdrucker/sportyR")
```

## Functional Overview

All functions in this library follow the same general format. They'll be called `geom_{sport}`, and take the following arguments:

 - `league`: the league for the sport. In all functions, this will ***NOT*** have a default value. Failure to supply will prompt the user for a league.
 
 - `full_surf`: a boolean indicating whether or not to plot the full surface. This defaults to `TRUE`. **NOTE**: this argument is not necessary for soccer pitches, baseball diamonds, or football fields, as these surfaces are always shown in full.
 
 - `rotate`: a boolean indicating whether or not to rotate the surface. All surfaces will be horizontal by nature. This defaults to `FALSE`
 
 - `rotation_dir`: the direction in which to rotate the plot (if `rotate == TRUE`). This defaults to `'ccw'` for counterclockwise

## Examples

Most playing surfaces are standard in size, so they can be rendered via a call to the proper `geom_{sport}` like so:

```{r nhl-example}
library(sportyR)
geom_hockey('nhl')
```

However, certain functions are able to take additional parameters. As an example, soccer pitches are not always a standard size. For this reason, users can specify in the call to `geom_soccer()` what the touchline and goal line dimensions should be (in meters).

```{r soccer-example-1}
geom_soccer('fifa')
```

```{r soccer-example-2}
geom_soccer('fifa', touchline_length = 100, goal_line_length = 75)
```
