---
title: "Customizing Plots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Customizing Plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
author: "Ross Drucker <br><a href='https://x.com/rossdrucker9' target='blank'><img src='https://img.shields.io/twitter/follow/rossdrucker9?color=blue&amp;label=%40rossdrucker9&amp;logo=x&amp;style=for-the-badge' alt='@rossdrucker9'/></a> <a href='https://github.com/rossdrucker' target='blank'><img src='https://img.shields.io/github/followers/rossdrucker?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@rossdrucker'/></a>"
description: "Recolouring, resizing and re-orienting a <code>sportyR</code> surface, overriding the league defaults to match a team palette or a different coordinate convention."
opengraph:
  image:
    src: "https://raw.githubusercontent.com/sportsdataverse/sportyR/main/man/figures/sportyR_gh.png"
  twitter:
    card: summary_large_image
    creator: "@sportyR_pkg"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "customize-plot-",
  out.width = "100%",
  dpi = 600
)
```

To run the following vignette, you'll need the `sportyR` package loaded into your workspace.

```{r setup}
library(sportyR)
```

Customizing a plot is easy as can be with `sportyR`. Say we want to customize a college basketball court to look like Lou Henson Court at the University of Illinois at Urbana-Champaign. You can see what it looks like [here](https://hoopdirt.com/wp-content/uploads/2015/08/louhensoncourt-rendering-690x340.jpg)

We won't get an exact match with logos, floorboard color differentiation, or some markings, but we'll get pretty close. Let's start by finding out what colors we can change, and how we can change them:

```{r customize-plot-get-colors}
# Find the colors for a college basketball court
cani_color_league_features("NCAA", "basketball")
```

Great, all set. It looks like the following should work:

***NOTE**: not all of the arguments below are needed, however all are shown to display the flexibility with which the plots can be customized.*

```{r customizing-plots-illinois-court, dev = "png"}
geom_basketball(
  league = "ncaa",
  color_updates = list(
    offensive_half_court = "#e8e0d7",
    defensive_half_court = "#e8e0d7",
    court_apron = "#e84a27",
    two_point_range = c("#e8e0d7", "#ffffff66"),
    center_circle_fill = "#e8e0d7",
    painted_area = c("#e84a27", "#ffffff00"),
    free_throw_circle_fill = "#e8e0d7",
    sideline = "#13294b",
    endline = "#13294b",
    division_line = "#13294b",
    center_circle_outline = "#13294b",
    lane_boundary = c("#ffffff", "#ffffff00"),
    three_point_line = c("#13294b", "#ffffff"),
    free_throw_circle_outline = "#ffffff",
    lane_space_mark = "#ffffff",
    restricted_arc = "#13294b",
    backboard = "#13294b"
  )
)
```

Pretty good! You can do this for any of the surfaces offered by `sportyR`. Here's another example, this time drawing a blue college football field:

```{r customizing-plots-blue-ncaaf-field, dev = "png"}
# Create a blue football field
geom_football(
  "ncaa",
  color_updates = list(
    field_apron = "#2e4597",
    field_border = "#2e4597",
    offensive_endzone = "#2e4597",
    defensive_endzone = "#2e4597",
    offensive_half = "#2e4597",
    defensive_half = "#2e4597",
    team_bench_area = "#2e4597"
  )
)
```

# **Our Authors**
-   [Ross Drucker](https://x.com/rossdrucker9)
<a href='https://x.com/rossdrucker9' target='blank'><img src='https://img.shields.io/twitter/follow/rossdrucker9?color=blue&amp;label=%40rossdrucker9&amp;logo=x&amp;style=for-the-badge' alt='@rossdrucker9'/></a>
<a href='https://github.com/rossdrucker' target='blank'><img src='https://img.shields.io/github/followers/rossdrucker?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@rossdrucker'/></a>

## **Citation**

To cite the [**`sportyR`**](https://sportyR.sportsdataverse.org/) R package in publications, use:

BibTeX Citation

```bibtex
@misc{sportyr,
  author = {Ross Drucker},
  title = {sportyR: Plot Scaled 'ggplot' Representations of Sports Playing Surfaces},
  url = {https://sportyR.sportsdataverse.org/},
  year = {2026}
}
```

## **Related SportsDataverse packages**

-   [**cfbfastR**](https://cfbfastR.sportsdataverse.org/) - college football
-   [**hoopR**](https://hoopR.sportsdataverse.org/) - men's basketball
-   [**wehoop**](https://wehoop.sportsdataverse.org/) - women's basketball
-   [**baseballr**](https://baseballr.sportsdataverse.org/) - baseball
-   [**fastRhockey**](https://fastRhockey.sportsdataverse.org/) - hockey
-   [**oddsapiR**](https://oddsapiR.sportsdataverse.org/) - betting odds
-   [**sportyR**](https://sportyR.sportsdataverse.org/) - playing surfaces
-   [**sportsdataverse-py**](https://py.sportsdataverse.org/) - the Python package
-   [**sportsdataverse-R**](https://r.sportsdataverse.org/) - the R meta-package
